---
title: "Exam 3"
author: "Rylie Butterfield"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Professor Salaries

First we want to load the libraries needed.
```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(ggplot2)
library(easystats)
library(broom)
```

Next, we want to load the data.
```{r message=FALSE, warning=FALSE}
sal = read.csv('FacultySalaries_1995.csv')
```

Then we want to clean the data.
```{r message=FALSE, warning=FALSE}

#pivot_longer() makes the data longer and puts all the ranks and salaries into their own columns.
#change the original column names 
sal2 = sal %>%
  pivot_longer(cols = c('AvgFullProfSalary', 'AvgAssocProfSalary', 'AvgAssistProfSalary'),
               names_to = 'Rank',
               values_to = 'Salary') %>%
  mutate(Rank = case_when(Rank == 'AvgFullProfSalary'~'Full',
                          Rank == 'AvgAssocProfSalary'~'Assoc',
                          Rank == 'AvgAssistProfSalary'~'Assist'))
```

Finally, we want to plot the data comparing Rank and Salary.
```{r echo=TRUE, message=FALSE, warning=FALSE}
sal2 %>%
  filter(Tier != 'VIIB') %>%
ggplot(aes(x = Rank, y = Salary, fill = Rank)) +
  geom_boxplot() +
  facet_wrap(~Tier) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

## Build and ANOVA Model
```{r echo=TRUE, message=FALSE, warning=FALSE}
#build the model
mod1 = aov(Salary ~ State + Tier + Rank, data = sal2)

#show the summary of the model
summary(mod1)
```


# Jupiter Oils
First we want to load the data.
```{r message=FALSE, warning=FALSE}
oil = read.csv('Juniper_Oils.csv')
```

Next, we want to clean the data.
```{r message=FALSE, warning=FALSE}
#Make the data longer and change the names in the ChemicalID column
oil2 = oil %>%
  pivot_longer(cols = 11:33,
               names_to = "ChemicalID",
               values_to = 'Concentration') %>%
  mutate(ChemicalID = case_when(ChemicalID =='alpha.pinene' ~ 'alpha-pinene',
                                ChemicalID =='para.cymene' ~ 'para-cymene',
                                ChemicalID =='alpha.terpineol' ~ 'alpha-terpineol',
                                ChemicalID =='cedr.9.ene' ~ 'cedr-9-ene',
                                ChemicalID =='alpha.cedrene' ~ 'alpha-cedrene',
                                ChemicalID =='beta.cedrene' ~ 'alpha-cedrene',
                                ChemicalID =='beta.cedrene' ~ 'beta-cedrene',
                                ChemicalID =='cis.thujopsene' ~ 'cis-thujopsene',
                                ChemicalID =='alpha.himachalene' ~ 'alpha-himachalene',
                                ChemicalID =='beta.chamigrene' ~ 'beta-chamigrene',
                                ChemicalID =='compound.1' ~ 'compound 1',
                                ChemicalID =='alpha.chamigrene' ~ 'alpha-chamigrene',
                                ChemicalID =='beta.acorenol' ~ 'beta-acorenol',
                                ChemicalID =='alpha.acorenol' ~ 'alpha-acorenol',
                                ChemicalID =='gamma.eudesmol' ~ 'gamma-audesmol',
                                ChemicalID =='beta.eudesmol' ~ 'beta-eudesmol',
                                ChemicalID =='alpha.eudesmol' ~ 'alpha-eudesmol',
                                ChemicalID =='cedr.8.en.13.ol' ~ 'cedr-8-en-13-ol',
                                ChemicalID =='cedr.8.en.15.ol' ~ 'cedr-8-en-15-ol',
                                ChemicalID =='compound.2' ~ 'compound 2',
                                ChemicalID =='cuparene' ~ 'cuparene',
                                ChemicalID =='widdrol' ~ 'widdrol',
                                ChemicalID =='cedrol' ~ 'cedrol',
                                ChemicalID =='thujopsenal' ~ 'thujopsenal',))
```

Then, we want to create plots for the concentration of each chemical in the years since the burn.
```{r echo=TRUE, message=FALSE, warning=FALSE}

oil2 %>%
  ggplot(aes(x = YearsSinceBurn, y = Concentration)) +
  geom_smooth() +
  facet_wrap(~ChemicalID, scales = 'free_y') +
  theme_minimal()
```

## Create a model to show which concentrations are significant

First, we need to create a model.
```{r echo=TRUE, message=FALSE, warning=FALSE}
#create the model
mod2 = glm(Concentration ~ YearsSinceBurn + ChemicalID, data = oil2)

#see information from the model
summary(mod2)
```

Next, we will limit the results to what concentrations are significant
```{r echo=TRUE, message=FALSE, warning=FALSE}
#choose what we want reported
sig_conc <- tidy(mod2) %>%
  filter(p.value < 0.05)

#show the results of what concentrations are significant
print(sig_conc)
```

